\speculake: Hiding Malware in Speculative Execution

### 2018-05-06 paper/architecture.tex
partial related work, and bib fixes across paper

\subsubsection{Reorder Buffer Capacity} \label{sssec:ROB}
We also measured the capacity of the reorder buffer (ROB) using a method
outlined by~\cite{intel-rob-capacity}. We measure the maximum number of cycles taken to
perform two uncached memory reads, and vary the number of filler instructions
between them. If the number of filler instructions is small, both memory reads
will fit inside the ROB, and can issue their memory reads in parallel. However,
if the filler instructions fill the ROB, the second memory read will have to
wait for the first to return before it can be issued, causing a noticeable step
increase in the cycle count.
Figure~\ref{fig:spec-capacity}
shows this step occurs at approximately 220 instructions for our processor,
suggesting a hard upper bound regardless of how long the cache miss takes to
resolve.

@@

We note a performance tradeoff between the size of communication (e.g. 4 bits vs 8
bits) and the time it takes the real world to recover the result from the side
channel. Using Flush+Reload~\cite{flush-reload} as our cache side channel,
recovering the result requires accessing all elements in an array exponential in
the size of the result (e.g. $2^8$ array reads to recover an 8-bit result). 
Therefore, there is a
performance advantage for keeping the size of the result small, and communicating
out small pieces of information that are aggregated by the real world.

>>>>

\subsubsection{Reorder Buffer Capacity} \label{sssec:ROB}
We also measured the capacity of the reorder buffer (ROB) using a method
outlined by~\cite{measuring-rob}. We measure the maximum number of cycles taken to
perform two uncached memory reads, and vary the number of filler instructions
between them. If the number of filler instructions is small, both memory reads
will fit inside the ROB, and can issue their memory reads in parallel. However,
if the filler instructions fill the ROB, the second memory read will have to
wait for the first to return before it can be issued, causing a noticeable step
increase in the cycle count.
Figure~\ref{fig:spec-capacity}
shows this step occurs at approximately 220 instructions for our processor,
suggesting a hard upper bound regardless of how long the cache miss takes to
resolve.

@@

We note a performance tradeoff between the size of communication (e.g. 4 bits vs 8
bits) and the time it takes the real world to recover the result from the side
channel. Using Flush+Reload~\cite{yarom2014flush+} as our cache side channel,
recovering the result requires accessing all elements in an array exponential in
the size of the result (e.g. $2^8$ array reads to recover an 8-bit result). 
Therefore, there is a
performance advantage for keeping the size of the result small, and communicating
out small pieces of information that are aggregated by the real world.

### 2018-05-06 paper/discussion.tex
partial related work, and bib fixes across paper

We note that this instruction seems to already exist, albeit unintentionally.
The \texttt{rep} instruction repeats the execution of string instructions a set
number of times (set in a register). When set to repeat 90 or more times, these
instructions are not speculated past, requiring they complete and be retired
before looking to the next instruction. This appears to be due to the 
\texttt{rep} instruction adding additional $\mu$ ops per instruction, we found
that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
operation works on $n$ bytes. Thus enough repetitions would cause instructions
after the \texttt{rep} instruction to wait for the full repetition to be retired
before beginning their execution.

>>>>

We note that this instruction seems to already exist, albeit unintentionally.
The \texttt{rep} instruction repeats the execution of string instructions a set
number of times (set in a register). When set to repeat 90 or more times, these
instructions are not speculated past, requiring they complete and be retired
before looking to the next instruction. This appears to be due to the 
\texttt{rep} instruction adding additional $\mu$ ops per instruction, we found
that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
operation works on $n$ bytes~\cite{uop_article}. Thus enough repetitions would cause instructions
after the \texttt{rep} instruction to wait for the full repetition to be retired
before beginning their execution.

### 2018-05-06 paper/related-work.tex
partial related work, and bib fixes across paper

This work makes use of ideas spanning multiple computer science and computer
security disciplines.

\subsection{CPU Reverse Engineering}
Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
multiple teams were working to reverse engineer the mechanisms that implemented in 
the microcode of various processors.

\subsection{Weird Machines}

\subsection{Red Pills}

@@

>>>>

% This work makes use of ideas spanning multiple computer science and computer
% security disciplines.

\subsection{Spectre}

% Lead in 
Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
multiple teams were working to reverse engineer the mechanisms that implemented in 
the microcode of various processors. As early as 2013 attempts to learn about the size 
of the re-order buffer were using cache timing side channels very similar to the 
Spectre variants~\cite{measuring-rob}.

% Discovery
Multiple teams discovered the Spectre vulnerabilites concurrently including the 
various teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
at Google~\cite{project_zero},and numerous others~\cite{}.

% Follow Up
The Spectre variants have been used in many other attacks on security critical 
including

% defenses 
Much like cache side channel vulnerabilities speculative execution is relatively
critical to processor performance and requires significant software, and hardware 
updates to fully address. Multiple measures have been proposed to address the 
individial variants of Spectre. Retpoline is Google's in house fix for enforcing 
isolation against branch target injection attacks~\cite{retpoline}. \textbf{Someone} 
\textit{et al.} propose that

We can certainly look forward to more vulnerabilities relating to speculative 
branching and related hardware side channels. Our work represents an exploit 
model for these vulnerabilities reliant upon shared hardware resources.

\subsection{Weird Machines \& Red-Pills}
Hardware side effects often leave behind unexpected state transitions 
that are not, or cannot be, accounted for in modeling environments. Malware
has historically made use of these bugs to create "weird" models of 
computation~\cite{weird_machines}. These state transitions can also leak 
information about the environment in which a program is running, which 
malware has used to avoid honey-pots and emulated environments.

The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
composing bugs and features of a system to create arbitrary compuatation with 
interesting properties~\cite{weird_machines,bratus2011exploit}. This 
encompasses many traditional computer vulnerabilities including 
buffer overflow~\cite{buffer_overflow}, format string exploits~\cite{format_string_exploit}, 
and return oriented programming~\cite{shacham2007geometry} to name only a few.

Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
can be constructed using the page fault handling mechanism in Intel's
IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
modern processors as multiple processes share system resources cooperatively
and optimizations and agregation are done objectively with regards of the 
process for the most part. Accessing copoerative hardware resources is at
the heart of the \speculake model, which constructs a hardware weird machine
using speculative branching and \texttt{Flush+Reload}. This is 
very much in the spirit of traditional exploit creation~\cite{weird_exploits}.

\smallskip

Similarly hardware bugs have been used to identify and fingerprint execution
environments, in malware this is know as a red-pill~\cite{red-pill}. This can be used 
to ensure that malware is not being emulated in a debugging 
environment~\cite{lindorfer2011detecting, balzarotti2010efficient}. Paleari 
\textit{et al.} automate the process of discovering red-pills by applying 
fuzzing techniques to emulated environmenta and identifying situations 
in which the emulators differ from their physical counterparts~\cite{paleari2009fistful}.
This is still by no means an exhaustive list of red-pills for a given 
processor as many rely on timing or race conditions in physical 
environments.

% Hardware bare-metal analysis
To combat this Kirat \textit{et al.} propose the use of automated "bare-metal" 
emulation of malware~\cite{kirat2011barebox}, and methods to determine if malware
is performing red-pill checks to avoid emulation environments~\cite{kirat2014barecloud}.
This can also be combatted using symbolic execution by finding environmental triggers 
in malicious binaries~\cite{schwartz2010all}.

\smallskip

Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
by patching binaries as the speculation can only happen on "bare-metal" systems, 
and break points will fail to trigger on a speculative call to a specific function.
Symbolic execution will also fail to find a target function as it is never on the 
execution pathway.

@@

The \speculake model relies critically on cache side channel attacks to 
communicate information from the speculative world.

\texttt{Flush+Reload}~\cite{yarom2014flush+}

Prime+Probe technique has been used against a number of processors
at multiple different leverls of cache. L1 data cache~\cite{} 
L2 cache~\cite{}
branch prediction cache~\cite{}
L1 instruction cache~\cite{}. These attacks have also been performed 
in cloud environments~\cite{}, collocated VMs~\cite{}, and local
environments~\cite{}.

While other works have demonstrated higher throughput from similar 
covert cache side channels~\cite{}, we incurr slightly higher performance 
penalties in our implementation as we perform redundancy checks to 
ensure the signal strength and cause intentional cache misses to enable
speculative execution costing hundreds to thousands of clock cycles~\ref{fig:cache-miss}
per round.
